The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-11 analysis/paper/paper.qmd
Move ngrams to pre-processing

```{r clean-text}
#| echo: false
#| dependson: get-pdf-text

emu$text_clean <- emu$text |>
  clean_basic()

@@

```{r stop-words}

# Add cleaned text column with removed stop words
emu$text_clean <- emu$text_clean |>
  str_remove_all(find_meta_stopwords(emu)) |> 
  str_remove_all(urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice", 
                      "space", "spatial", "spaces", "plan", "public", "development"))) |>
  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning")))

# Unnest words and remove words of one and two letters
emu_words <- emu |> 
  select(title, text_clean) |> 
  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
  mutate(word = textstem::lemmatize_words(word))  # lemmatise words

@@

```{r short-stop-words}
#| dependson: stop-words

#short_words(emu_words$word, l_word = 3, l_list = 100)

emu_words <- emu_words |>  
  filter(nchar(word) >= 3)  # keep only words that are at least three letters long

@@

```{r topic-model-lda-1}
#| dependson: short-stop-words

@@

# Reduce the size of the DTM by excluding words with a frequency lower than 5
emu_dtm <- convert_to_dtm(emu_theses_words, 
                          title_col = "title", 
                          word_col = "word", 
                          min_term_freq = 5)

@@

```{r bigrams}

emu_bigrams <- emu |> 
  get_ngrams(n = 2, title_col = "title", text_col = "text_clean", lemma = TRUE) |> 
  unite(ngram_lemma, w_1_lemma, w_2_lemma, sep = " ", remove = FALSE)

bigram_counts <- emu_bigrams |> 
  count(ngram, sort = TRUE) |> 
  filter(n > 20)

emu_bigrams <- emu_bigrams |> 
  filter(ngram %in% bigram_counts$ngram)

bigram_lemma_counts <- emu_bigrams |> 
  count(w_1_lemma, w_2_lemma, sort = TRUE) |> 
  filter(n > 5)

# Concatenate bigrams

# # First attempt using a named vector in `str_replace_all()`
c_bigrams <- paste(emu_bigrams$ngram_lemma, sep = "_")
names(c_bigrams) <- emu_bigrams$ngram

system.time(
  emu$text_clean <- emu$text_clean |>
    str_replace_all(fixed(c_bigrams))
)

# Second attempt with `mgsub::mgsub()`
# emu$text_clean <- mgsub(emu$text_clean,
#                         pattern = emu_bigrams$ngram,
#                         replacement = emu_bigrams$ngram_lemma)

```{r trigrams}

## Trigrams
emu_trigrams <- emu |> 
  get_ngrams(n = 3, title_col = "title", text_col = "text_clean", lemma = TRUE)

trigram_counts <- emu_trigrams |> 
  count(w_1, w_2, w_3, sort = TRUE)

trigram_lemma_counts <- emu_trigrams |> 
  count(w_1_lemma, w_2_lemma, w_3_lemma, sort = TRUE)

>>>>

@@

```{r pre-process-text}

meta_sw <- find_meta_stopwords(emu)
urbanism_sw <- urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism", 
                      "european post master in urbanism",
                      "european postmaster in urbanism",
                      "tu delft", "ku leuven", "upc barcelona", "iuav venice"))
thesis_sw <- thesis_stopwords(add_stopwords = c("advisor", "prof", "fig", "ure", "ning"))
custom_sw <- c("space", "spatial", "spaces", "plan", "public", "development") |>
  (\(.) paste0("\\b", ., "\\b"))() |> 
  paste(collapse = "|")

# Add cleaned text column
emu$text_clean <- emu$text_raw |>
  # Remove white spaces and punctuation, and change text to lower-case
  clean_basic() |>
  # Remove title, subtitle and full title
  str_remove_all(meta_sw) |> 
  # Remove dataset-specific words
  str_remove_all(urbanism_sw) |>
  # Remove thesis-specific words
  str_remove_all(thesis_sw)

# Concatenate ngrams
emu$text_clean <- emu |> 
  c_ngrams(n_max = 3, text_col = "text_clean",
           title_col = "title", min_freq = 20, lemma = TRUE)

  # Words that were not included in ngrams are safe to remove now
emu$text_clean <- emu$text_clean |> 
  str_remove_all(custom_sw)

```{r tokenise}

# TODO Replace title column with ID column

# Unnest tokens and remove words of one and two letters
emu_words <- emu |> 
  select(title, text_clean) |> 
  unnest_tokens(output = word, input = text_clean) |>  # remove punctuation, convert to lower-case, separate all words
  anti_join(tidytext::stop_words, by = "word") |>  # remove stop words
  mutate(word = textstem::lemmatize_words(word)) |>  # lemmatise words
  filter(nchar(word) >= 3)  # keep only words that are at least three letters long

@@

@@

```{r topic-model-lda-1}

@@

# Reduce the size of the DTM by excluding words with a frequency lower than 5
emu_dtm <- convert_to_dtm(emu_words, 
                          title_col = "title", 
                          word_col = "word", 
                          min_term_freq = 5)

@@

A frequency count of ngrams reveals collocations like "public space" or "socio-spatial segregation" across the corpus. To include these in the topic model, we concatenated them with an underscore, so that they are considered one word during the word-level tokenisation. Only bigrams and trigrams, that is two- and three-word collocations, were extracted from the corpus.
